\documentclass[times,11pt]{article}
\usepackage{latexsym}
\renewcommand{\baselinestretch}{0.9}
\oddsidemargin -0.15in     % Real margin = (210 - 30) / 144 = 1.25 in
\evensidemargin -0.15in
\topmargin -0.2in       % Real topmargin = (140 - 68) / 144 = 0.5 in
\textheight 9.0in    % 626 pt = 8.66 in
\textwidth 6.5in     % 434 pt = 6 in
\reversemarginpar

\newcommand{\fijisys}{Fiji Systems Inc.}


\begin{document}


\onecolumn

\pagestyle{empty}

\bigskip
\begin{center}
{\large {\bf{Lukasz Ziarek}}}\\
\vspace{0.3em}
338E Davis Hall, University at Buffalo, Buffalo, NY 14260-2500
\end{center}

\newlength{\myspace}
\setlength{\myspace}{-.5mm}%-.4mm}

\subsection*{Professional Preparation}


\begin{tabular}{p{8cm}@{\hspace{2mm}}l@{\hspace{5mm}}l}
University of Chicago & Computer Science & B.S.     
December 2003 \\
Purdue University & Computer Science & Ph.D.       
May 2011 \\
\end{tabular}

\subsection*{Appointments}
\noindent\begin{tabular}{ll}
Assistant Professor,   Department of Computer Science and Engineering, SUNY Buffalo. & 2012 -- \\
President, \fijisys, South Bend, IN. & 2010 -- \\
Visiting Assistant Professor, Department of Computer Science, Purdue University. & 2011 -- 12\\
Vice-President, \fijisys, Indianapolis, IN. & 2009 -- 10\\
Research Assistant, Department of Computer Science, Purdue University.  & 2004 -- 11\\
\end{tabular}

\subsection*{Products}

\noindent
\begin{enumerate}\setlength{\itemsep}{\myspace}

\item KC Sivaramakrishnan, Lukasz Ziarek, Suresh Jagannathan:
{\bf A Coherent and Managed Runtime for ML on the SCC.}
{\it Many-core Applications Research Community Symposium} 2012.

\item Lukasz Ziarek, KC Sivaramakrishnan, and Suresh Jagannathan:
{\bf Composable Asynchronous Events.}
In \emph{Programming Language Design and Implementation} 2011.

\item Lukasz Ziarek and Suresh Jagannathan:
{\bf Lightweight Checkpointing for Concurrent ML.} 
In \emph{ Journal of Functional Programming} 2010.

\item  Lukasz Ziarek, KC Sivaramakrishnan, and Suresh Jagannathan:
{\bf Partial Memoization of Concurrency and Communication.} 
In \emph{International Conference on Functional Programming} 2009.

\item Lukasz Ziarek, Adam Welc, Ali-Reza Adl-Tabatabai, Vijay Menon, 
Tatiana Shpeisman, and Suresh Jagannathan:
{\bf A Uniform Transactional Execution Environment for Java.} 
In \emph{European Conference on Object-Oriented Programming} 2008.  

\end{enumerate}


\subsection*{Significant Products}

\noindent
\begin{enumerate}\setlength{\itemsep}{\myspace}

\item Ethan Blanton, Demian Lessa, Lukasz Ziarek, and Bharat Jayaraman:                                                                                                 
{\bf JI.FI: Visual Test and Debug Queries for Hard Real-Time.}                                                                                                          
{\it International Workshop on Java Technologies for Real-time and Embedded Systems} 2012.  

\item Lukasz Ziarek, Siddharth Tiwary, and Suresh Jagannathan:
{\bf Isolating Determinism in Multi-Threaded Programs.}
In \emph{Runtime Verification} 2011.

\item Filip Pizlo, Lukasz Ziarek, Ethan Blanton, Petr Maj and Jan Vitek:
{\bf High-level Programming of Embedded Hard Real-Time Devices.}
In \emph{EuroSys} 2010.

\item Filip Pizlo, Lukasz Ziarek, Petr Maj, Anthony Hosking, Ethan Blanton, and Jan Vitek:
{\bf Schism: Fragmentation-Tolerant Real-Time Garbage Collection.}
In \emph{Programming Language Design and Implementation} 2010.

\item Lukasz Ziarek, Stephen Weeks, and Suresh Jagannathan:
{\bf Flattening Tuples in an SSA Intermediate Representation.}
In \emph{Higher Order and Symbolic Computation, Volume 23, Number 3} 2008.

\end{enumerate}


\subsection*{Synergistic Activities}
\noindent
\begin{enumerate}\setlength{\itemsep}{\myspace}
\item Outreach: Member of the Computer Science Teacher's Association.
\item Awards: Halstead Award for Outstanding Research in Software Engineering, 2009;
Intel Fellowship, 2008; 
Department Of Education Graduate Assistance In Areas Of National Need Fellowship, 2004.
\item Program Committees: Java Technologies for Real-time and Embedded Systems, 2010, 2011; 
Declarative Aspects for Multi-Core Programming, 2012.
\item Journal Referee: Concurrency and Computation Practice and Experience; Software: Practice and Experience.
\end{enumerate}


\subsubsection*{(Collaborators \& Other Affiliations)}
\subsubsection*{Collaborators and Co-Editors}
\noindent Umut Acar (CMU),
Ali-Reza Adl-Tabatabai (Intel), 
Ethan Blanton (Purdue University, \fijisys),
Patrick Eugster (Purdue University),
Mathew Fluet (Rochester), 
Anthony Hosking (Purdue University),
Suresh Jagannathan (Purdue University),
Bharat Jayaraman (SUNY Buffalo), 
Demian Lessa (SUNY Buffalo),
Petr Maj (Purdue University),
Vijay Menon (Google),
Filip Pizlo (Apple Inc.),
Raghavendra Prasad (Google),
Tatiana Shpeisman (Intel),
KC Sivaramakrishnan (Purdue University),
Jan Vitek (Purdue University, \fijisys),
Adam Welc (Oracle)


\subsubsection*{Graduate Advisors and Postdoctoral Sponsors}
\noindent \begin{tabular}{lcc}
Suresh Jagannathan & Purdue University & West Lafayette, Indiana, USA.\\
Patrick Eugster    & Purdue University & West Lafayette, Indiana, USA.\\
Zihuan Li          & Purdue University & West Lafayette, Indiana, USA.\\
Jan Vitek          & Purdue University & West Lafayette, Indiana, USA.\\
\end{tabular}

\subsubsection*{Thesis Advisor and Postgraduate-Scholar Sponsor}

\textbf{Thesis Advisor:} (3) \\
\emph{Graduated:} none\\
\emph{Current:} Shaun Cosgrove, Feng Shen, Yin Yan\\
\textbf{Postgraduate-Scholar Sponsor:} (0) \\
none

\end{document}
%\subsubsection*{Conference Organization (Program Committee Memberships unless specified)}
%Java Technologies for Real-time and Embedded Systems, 2010.
%Java Technologies for Real-time and Embedded Systems, 2011.
%Declarative Aspects for Multi-Core Programming, 2012.



